---
title: 'Statistical report: Assessment of preceding heart failure admission as a predictor of outcomes in TOPCAT'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

TOPCAT

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

## Fixes

Values < 0 in the continuous variables used in the models are replaced by missing (BARNA: Inte bara HB). 

## Frågor/Kommentarer

Kodade ihop tertilerna av NT-proBNP och BNP till en variabel som Lars föreslog någon gång. De blir ändå en hel del mindre missing. OK?

Prev HF admission inkluderar även om de är randomiserade in hosp. Behöver detta dock vara just HF hosp? Skriva i limitaitons?

Fler variabler for justering i modeller?

I TOPCAT verkar de ha kodat Unknown död som EJ cvd. Drog till med sens analys där kodade Unknown som ej cvd (dock bara för primärt utfall, kan göra även för andra), men kan lyfta bort om blir för mycket info. 

Det finns inte discharge date i adj events filen. Och det går inte att matcha ihop med t027 (sitens rapporterade events) för datumen diffar. Därför får time at risk vara även när pat är inneliggande för hosp. 

Har 3 previous HDH variabler och 5 utfall. Det blir många olika kombos. 
Går eliminera någon? Bör välja primär HFH och utfall. Har valt previous HFH inom 1 år som primär variabel (eftersom spline analyserna är på denna). 
vet dock inte vad som är nytt och om riskerar att inte se det nya för alla andra siffror. Kan kanske också nöja sig med att presentera sens analyser (unknown death as non-cvd, exkludera ntprobnp/bnp för justering mm ENBART för primary outcome). Ska CVD/rec HFH vara primärt utfall (fördel: brukar vara det i trials nackdel: effekten av prev HFH borde vara främst på HFH, så man vill väll egentligen helst se de sep utfallen cvd och hfh var för sig?).  

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with median for continuous variables and mode for categorical variables. 
For the combined ntprobnp/bnp tertile variable missing values were coded as a separate category. 
A sensitivity analyses was also performed excluding the NT-proBNP/BNP variable.

## Previous HFH

All analyses are performed for 3 different sets of previous HFH

- Previous HFH within 1 year from randomisation vs no previous HFH within 1 year (primary exposure)
- Previous HFH vs no previous HFH
- No previous HFH, HFH < 30d, HFH 30-90d, HFH 90d-6mo, HFH 6mo-1yr, HFH > 1yr. (FÄRRE GRUPPER, ANNAN INDELNING?)

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Associations between previous HFH and outcomes

The following outcomes are considered: 

- CV mortality + total HF hospitalization (primary endpoint)
- CV mortality
- Total HF hospitalization
- CV mortality + first HF hospitalization
- First HF hospitalization

Time to first events were presented with cumulative incidence curves and 
the mean cumulative function (MCF) was used to depict the recurrent events for previous HFH within 1 year. 
SÄG TILL OM NI VILL HA FÖR DE ANDRA HFH VARIABLERNA OCKSÅ. 

Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only previous HFH in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab11yr}. The variables were selected based on clinical relevance. 
The continuous variables were included using natural cubic splines with 5 degrees of freedom. 
Due to violations to the proportional hazards assumption loop diuretics was stratified for in the models.  

As a consistency analysis the time to first events were modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

Recurrent events were model using the Prentice-Williams-Peterson total-time model, crude and adjusted as above. 
As a consistency analysis the recurrent events were modelled using a negative binomial regression 
including the log of time as an offset in the model. 

The median (min-max) follow-up is 
`r tcdata %>% summarise(med = fn(median(outtime_cvd), dig = 1),
                                             min = fn(min(outtime_cvd), dig = 1),
                                             max = fn(max(outtime_cvd), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r tcdata %>% 
                                   summarise(sumpy = fn(sum(outtime_cvd), dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.
                                   
\clearpage

```{r, child = "./src/km.Rmd"}

```

\clearpage

```{r, child = "./src/outtab.Rmd"}

```

### Visualization of association between previous HFH within 1 year and outcomes

Previous HFH on a continuous scale was visually presented using natural cubic splines with 5 degrees of freedom. 
The model only included patients with a previous HFH within 1 year from randomization. 

```{r, child = "./src/splines.Rmd"}

```

\clearpage

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. Loop diuretics was as a results including as a strata variable.  
Possible outliers were visually inspected by plotting the dfbetas. No action deemed necessary. 
Since age, map and bmi showed signs of non-linear associations with the outcome all all variables were including using natural cubic splines. 

### Subgroups

Performed using an interaction term between previous HFH and the selected respective variables. 

WHICH SUBGROUPS???? AND FOR WHAT OUTCOME/previous HFH variable?

```{r, child = "./src/outforestsub.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/topcat_hfhosp. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References