---
title: 'Statistical report: Assessment of preceding heart failure admission as a predictor of outcomes in TOPCAT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

TOPCAT

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

## Fixes

Values < 0 in the continuous variables presented are replaced by missing. Time from previous HFH that is after randomization is set to 0. 

NT-proBNP and BNP are categorized according to tertiles and the tertiles are combined into one variable. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with median for continuous variables and mode for categorical variables. 

For the combined ntprobnp/bnp tertile variable missing values were coded as a separate category. 
A sensitivity analysis was also performed without adjustment for NT-proBNP/BNP in the model.

Unknown causes of death are assumed to be CV. A sensitivity analysis was 
performed were unknown was assumed to be non-CV.

## Previous HFH

All analyses are performed for:

- No previous HFH, in-hospital, HFH 1-30d, HFH 31-90d, HFH 91-180d, HFH 181-365d, HFH > 365d
- Previous HFH vs no previous HFH (at any time point)

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Associations between previous HFH and outcomes

The following outcomes are considered: 

- CV mortality + total HF hospitalization (primary endpoint)
- CV mortality
- Total HF hospitalization
- CV mortality + first HF hospitalization
- First HF hospitalization

Time to first events were presented with cumulative incidence curves (CIC) and 
the mean cumulative function (MCF) was used to depict the recurrent events by previous HFH. 

Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only previous HFH in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab1cat}. The variables were selected based on clinical relevance. 
The continuous variables were included using natural cubic splines with 4 degrees of freedom. 
Due to violations to the proportional hazards assumption loop diuretics was stratified for in the models.  

As a consistency analysis the time to first events were modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

Recurrent events were model using the Prentice-Williams-Peterson total-time model, crude and adjusted as above. 
As a consistency analysis the recurrent events were modelled using a negative binomial regression 
including the log of time as an offset in the model. 

Time was censored at end of study or death not defined as an outcome. 

The median (min-max) follow-up is 
`r tcdata %>% summarise(med = fn(median(outtime_cvd), dig = 1),
                                             min = fn(min(outtime_cvd), dig = 1),
                                             max = fn(max(outtime_cvd), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r tcdata %>% 
                                   summarise(sumpy = fn(sum(outtime_cvd), dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.
                                   
\clearpage

```{r, child = "./src/km.Rmd"}

```

\clearpage

```{r, child = "./src/outtab.Rmd"}

```

### Visualization of association between time from previous HFH and outcomes

Previous HFH on a continuous scale was visually presented using natural cubic 
splines with 4 degrees of freedom using the
Cox proportional hazards regressions adjusted for the same variables as above for time to first event and
the Prentice-Williams-Peterson total-time model, adjusted as above, for recurrent events. 
The reference (HR = 1) in the plots is the median time from previous HFH.  

The model only included patients with a previous HFH. 

```{r, child = "./src/splines.Rmd"}

```

\clearpage

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. Loop diuretics was as a results including as a strata variable.  
Possible outliers were visually inspected by plotting the dfbetas. No action deemed necessary. 
Since age, map and bmi showed signs of non-linear associations with the outcome all all variables were including using natural cubic splines. 

### Subgroups

Performed using an interaction term between previous HFH and the selected 
respective variables for the outcome CV mortality + total HF hospitalization (primary endpoint). 

```{r, child = "./src/outforestsub.Rmd"}

```

\clearpage
\newpage

### Association between Spironolactone and outcomes in different previous HFH subgroups

Performed using an interaction term between Spironolactone and previous HFH groups for the outcome CV mortality + total HF hospitalization (primary endpoint). Adjustment is not performed. 

```{r, child = "./src/outforestsubspiro.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/topcat_hfhosp. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
