---
title: 'Statistical report: Assessment of preceding heart failure admission as a predictor of outcomes in TOPCAT'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

TOPCAT

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

## Frågor/Kommentarer

Kodade ihop tertilerna av NT-proBNP och BNP till en variabel som Lars föreslog någon gång. De blir ändå en hel del mindre missing. OK?

Prev HF admission inkluderar även om de är randomiserade in hosp. Behöver detta dock vara just HF hosp? Skriva i limitaitons?

32 personer i TOPCAT americas uppfyllde inte inklusions och exklusions kriterierna. I ditt datahanteringsdokument Barna står att dessa är exkluderade tror inte du gjorde det till slut ändå (då stämmer inte antalen). Vanligtvis (så länge de är randomiserade) så inkluderas dessa i ITT/FAS population (primär population) så antar att dessa ska inkluderas i denna studie. Dessa är också med i TOPCAT pek. OK?

Jag exkluderade även 2 personer med saknad CHF_HOSP jmf med dig Barna. OK? 

Jag använde MEDCAT_WHODDE istället för att koda ihop medicinerna. Förutom COMBO gjorde jag inga omkodningar. Det diffar därför lite mellan våra meds siffror. 

Fler variabler for justering i modeller?

För utfallen tog jag med de fall som skett efter study end date (till skillnad från dig Barna). Det är så de verkar ha gjort i TOPCAT (dataset Outcomes) så föreslår gör likadant. 

I TOPCAT verkar de ha kodat Unknown död som EJ cvd. Drog till med sens analys där kodade Unknown som ej cvd, 
men kan lyfta bort om blir för mycket info. 

Du skriver att du använt T027 för hfh. Tror ska använda T070? Det är det de verkar ha gjort i TOPCAT. Som jag uppfattar det är T027 hosp rapporterade, och T070 är adjudicerade. Eller missförstår jag något? Sen förstår jag inte riktigt hur det gjort i TOPCAT (de har i vissa fall tagit en senare event istället för första (om outcome nummret är lägre), och för cvd har de i första hand använd cea datumet, för hfh i första hand site datumet. Jag gör likadant som de, men misstänker att de nog gjort det aningens småknasigt. 

Det finns inte discharge date i adj events filen. Och det går inte att matcha ihop med t027 (sitens rapporterade events) för datumen diffar. Därför får time at risk vara även när pat är inneliggande för hosp. 

HUVUDFRÅGA: Har 3 previous HDH variabler och 5 utfall. Det blir många olika kombos. 
Går eliminera någon? Bör välja primär HFH och utfall. Har valt previous HFH inom 1 år som primär variabel (eftersom spline analyserna är på denna). 
vet dock inte vad som är nytt och om riskerar att inte se det nya för alla andra siffror. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with median for continuous variables and mode for categorical variables. 
For the combined ntprobnp/bnp tertile variable missing values were imputed with Tertile 2. 
A sensitivity analyses was also performed excluding the NT-proBNP/BNP variable HAR INTE GJORT DETTA ÄR SÅ MÅNGA ANALYSER SOM DET ÄR SÅ FÖRESLÅR SKIPPA. 

## Previous HFH

All analyses are performed for 3 different sets of previous HFH

- Previous HFH within 1 year from randomisation vs no previous HFH within 1 year (primary exposure)
- Previous HFH vs no previous HFH
- No previous HFH, HFH < 30d, HFH 30-90d, HFH 90d-6mo, HFH 6mo-1yr, HFH > 1yr. (FÄRRE GRUPPER, ANNAN INDELNING?)

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Associations between previous HFH and outcomes

The following outcomes are considered: 

- CV mortality + recurrent HF hospitalization (primary endpoint)
- CV mortality
- Recurrent HF hospitalization
- CV mortality + first HF hospitalization
- First HF hospitalization

Time to first events were presented with cumulative incidence curves and 
the mean cumulative function (MCF) was used to depict the recurrent events for previous HFH within 1 year. 
SÄG TILL OM NI VILL HA FÖR DE ANDRA HFH VARIABLERNA OCKSÅ. 

Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only previous HFH in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab11yr}. The variables were selected based on clinical relevance. 
As a consistency analysis the time to first events were modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

Recurrent events were model using the Prentice-Williams-Peterson total-time model, crude and adjusted as above. 
As a consistency analysis the recurrent events were modelled using a negative binomial regression 
including the log of time as an offset in the model. 

The median (min-max) follow-up is 
`r tcdata %>% summarise(med = fn(median(outtime_cvd), dig = 1),
                                             min = fn(min(outtime_cvd), dig = 1),
                                             max = fn(max(outtime_cvd), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r tcdata %>% 
                                   summarise(sumpy = fn(sum(outtime_cvd), dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.
                                   
\clearpage

```{r, child = "./src/km.Rmd"}

```

\clearpage

```{r, child = "./src/outtab.Rmd"}

```

### Visualization of association between previous HFH and outcomes

Previous HFH on a continuous scale was visually presented using natural cubic splines with 5 degrees of freedom. 
The model only included patients with a previous HFH within 1 year from randomization. 

INTE SNYGGA GRAFER, SKA FÖRSÖKA FIXA TILL. 

```{r, child = "./src/splines.Rmd"}

```

\clearpage

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> XXXX. Linearity..... CHECK THIS LATER. 

### Subgroups

Performed using an interaction term between previous HFH and the respective variables. 

WHICH SUBGROUPS???? AND FOR WHAT OUTCOME/previous HFH variable?

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/topcat_hfhosp. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References